Polynomial phase estimation by phase unwrapping

### 2012-10-16 paper.tex
minor edits.

% put your definitions there:
\startlocaldefs
\newcommand{\reals}{{\mathbb R}}
\newcommand{\ints}{{\mathbb Z}}
\newcommand{\term}{\emph}
\newcommand{\var}{\operatorname{var}}
\newcommand{\covar}{\operatorname{covar}}
%\newcommand{\prob}{{\mathbb P}}
\newcommand{\prob}{\operatorname{Pr}}
\newcommand{\expect}{{\mathbb E}}
\newcommand{\dealias}{\operatorname{dealias}}
\renewcommand{\mid}{\; ; \;}
\newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
\newcommand{\sfracpart}[1]{\langle #1 \rangle}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\sabs}[1]{\vert #1 \vert}
\newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
\newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
\usepackage{mathbf-abbrevs}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newcommand{\coef}{\operatorname{coef}}
\newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
\newcommand{\scubr}[1]{{\{ #1 \}}}
\newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
\endlocaldefs

@@

An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $N^{-k}$.  This is perhaps not surprising, since it is the same rate observed in polynomial regression.  However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.  For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$.  %This proof technique appears to be novel, and may be useful in other problems.  
The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.  Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.  We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.  Potential candidates are the phase wrapped images observed in synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.

@@

It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice. Polynomial time algorithms that compute the nearest point are described in~\cite[Sec 4.3]{McKilliam2010thesis}.  Although polynomial in complexity, these algorithms are not fast in practice.  The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.  In this paper we focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects.

@@

 %\begin{equation}
 %E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=\sum_{n_{1},\ldots,n_{2\beta
 %}=1}^{N}E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)  .
 %\label{eq:markapp}%
 %\end{equation}
 %Since the terms%
 %\[
 %E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)
 %\]
 %in this containing any $n_{j}$ only once will have $0$ mean, and since the
 %$z_{j}$ are bounded, it follows that%
 %\[
 %\frac{1}{N^{\beta}}E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=O\left(
 %1\right)  ,
 %\]
 %as $N\rightarrow\infty,$ the dominant terms coming from the $z_{j}$ in
 %$\left(  \ref{eq:markapp}\right)  $ occurring in pairs.

@@

This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU). It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed. %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.  
It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.  Polynomial time algorithms that compute the nearest point are described in~\cite[Sec 4.3]{McKilliam2010thesis}, but these are slow algorithms in practice.
A significant outstanding question is whether practically fast nearest point algorithms exist for these specific lattices.  Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.

@@

During the proof of strong consistency we made use of the fact that
\begin{equation}~\label{eq:SNVNunifmlln2}
\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
\end{equation}
almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here. Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.  By Lemma~\ref{lem:vn},
\[
\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty
\]
for any $\epsilon > 0$.  Thus~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.  In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.

\begin{lemma} \label{lem:vn} 
For any $\epsilon > 0$ and $0 < c < 8$,
\[
\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
\]
 \end{lemma}
\begin{proof}
Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
 %\begin{align*}
\[ 
%\lambdabf[\rbf] = &\left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \right.\\ 
%&\dots, \; \frac{r_k}{N^{b+k}} - \frac{1}{2(k!)}, \; \dots \\
%&\left. \dots, \; \frac{r_m}{N^{b+m}} - \frac{1}{2(m!) }\right]
 \lambdabf[\rbf] = \left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \; \dots, \; \frac{r_m}{m!N^{b+m}} - \frac{1}{2(m!) }\right]
\]
 %\end{align*}
 for some constant $b>0$.  Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the first coordinate, $\tfrac{1}{N^{b+1}}$ in the second coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate. Let
 \[
 B[\rbf]=\left\{  \xbf\in\reals^{m+1}  ; \frac{r_k}{N^{b+k}}\leq x_k + \frac{1}{2(k!)} < \frac{r_k + 1}{N^{b+k}} \right\}  .
 \]
and let $G$ be the finite set of grid points
\[
G = \left\{ \xbf \in \ints^{m+1} \mid x_k = 0,1,2\dots,N^{b+k}-1  \right\}.
\]
The total number of grid points is $|G| = N^{(m+1)(2b + m)/2}$, and the $B[\rbf]$ partition $B$, that is, $B = \cup_{\rbf \in G}B[\rbf]$.  Now 
\begin{align}
 \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
 &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
 \end{align}
From Lemma~\ref{lem:supVjk} it will follow that
\[
\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
\]
for any real number $c < 2$.
%that is, with high probability $D_N$ is small on all of the grid points.  
In Lemma~\ref{lem:supBVn} we show that
\[
\sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\abs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])} < 2\frac{m+1}{N^b}.
\]
%that is, $D_N$ does not change much within any of the regions $B[\rbf]$ neighbouring the grid points.  
Combining these results with~\eqref{eq:gridanddensespace}, we obtain
\[
\prob\left(\sup_{\lambdabf \in B }\sabs{ D_N(\lambdabf)} > \frac{\epsilon}{2} +  \frac{2(m+1)}{N^b} \right) = O(e^{-c \epsilon^2 N}),
\]
and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
\end{proof}

@@

\begin{lemma}\label{lem:supVjk}
For any $\epsilon > 0$ and $0 < c < 8$,
\[
\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N}).
\]
\end{lemma}
 \begin{proof}
Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
 \[
 Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - E\fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
 \]
 are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
 \begin{equation}
  \prob( \abs{D_N(\lambdabf)} > \epsilon ) \leq 2e^{ - 8 \epsilon^2 N}.
 \label{eq:markapp}%
 \end{equation}
But,
 \begin{align*}
 \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
&= 2 |G| e^{ - 8 \epsilon^2 N} = O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
\end{align*}
where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
\end{proof}

@@

\begin{lemma}\label{lem:boundedsquarefracparts}
Let $x$ and $\delta$ be real numbers.  Then
\[
\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|.
\]
\end{lemma}
\begin{proof}
Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Also, for reasons of symmetry, we need only show that it holds when $\delta \geq 0$.  Now
\[
\fracpart{x + \delta}^2 - x^2 = \begin{cases}
2x\delta + \delta^2, & x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta) \\
2x(\delta-1) + (\delta-1)^2, & x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})
\end{cases}
\] 
But, when $x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta)$, 
\[
-1 \leq -1 + \delta \leq 2x + \delta < 1 - \delta \leq 1,
\]
and so
\[
-\delta \leq (-1 + \delta)\delta \leq (2x + \delta)\delta < (1-\delta)\delta \leq \delta.
\]
Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$,
\[
-\delta \leq 2x + \delta - 1 < \delta,
\]
and consequently
\[
-\delta \leq -\delta(1-\delta) \leq (2x + \delta - 1)(1 - \delta) \leq \delta(1 - \delta) \leq \delta.
\]
%and consequenly
%\[
%-\delta \leq (2x + \delta - 1)(\delta - 1) \leq \delta.
%\]

@@

 \begin{lemma}\label{lem:supBVn} $\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}$ for all $N$.
 \end{lemma}
 \begin{proof}
%The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
For convenience put 
\[
b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \qquad \text{and} \qquad  a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
\] 
where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
where 
\[
\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
\] 
From Lemma~\ref{lem:boundedsquarefracparts} it follows that $-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|$,
and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all real numbers $x$. Now
\[
S_N(\lambdabf) - S_N(\lambdabf[\rbf]) =  \frac{1}{N}\sum_{n=1}^{N}\big( \fracpart{\Phi_n + b_n}^2 -  \fracpart{\Phi_n + a_n}^2 \big)
\]
and therefore, for all $\lambdabf \in B[\rbf]$,
\[
\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}.
\]
As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
\[
\sabs{ V_N(\lambdabf) - V_N(\lambdabf[\rbf])} \leq \expect \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
\]
by Jensen's inequality.  Therefore, for all $\lambdabf \in B[\rbf]$,
 \begin{align*}
 \sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } &= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
&\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
&\leq 2\frac{m+1}{N^{b}},
\end{align*}
and the lemma follows because this bound is independent of $\rbf$,
%\[
%\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\sabs{ D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right) } \leq 2\frac{m+1}{N^{b}}.
%\]
 \end{proof}

@@

During the proof of asymptotic normality in Lemma~\ref{lem:Kconvfhalf} we made use of the following result regarding the function,
\[
G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
\]
where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.  To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.  The proofs we will give hold for any nonegative integer $\ell$.

@@

 \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \\

>>>>

% put your definitions there:
\startlocaldefs
\newcommand{\reals}{{\mathbb R}}
\newcommand{\ints}{{\mathbb Z}}
\newcommand{\term}{\emph}
\newcommand{\var}{\operatorname{var}}
\newcommand{\covar}{\operatorname{covar}}
%\newcommand{\prob}{{\mathbb P}}
\newcommand{\prob}{\operatorname{Pr}}
\newcommand{\expect}{{\mathbb E}}
\newcommand{\dealias}{\operatorname{dealias}}
\renewcommand{\mid}{\; ; \;}
\newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
\newcommand{\sfracpart}[1]{\langle #1 \rangle}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\sabs}[1]{\vert #1 \vert}
\newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
\newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
\usepackage{mathbf-abbrevs}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newcommand{\coef}{\operatorname{coef}}
\newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
\newcommand{\scubr}[1]{{\{ #1 \}}}
\newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
\newcommand{\bfracpart}[1]{\big\langle #1 \big\rangle}
\newcommand{\babs}[1]{\big\vert #1 \big\vert}
\endlocaldefs

@@

An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $N^{-k}$.  This is perhaps not surprising, since it is the same rate observed in polynomial regression.  However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.  For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$.  %This proof technique appears to be novel, and may be useful in other problems.  
The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.  Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.  We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.  Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.

@@

It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice. Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}.  Although polynomial in complexity, these algorithms are not fast in practice.  The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.  In this paper we focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects.

@@

@@

This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU). It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed. %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.  
It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.  Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}, but these are slow algorithms in practice.  A significant outstanding question is whether practically fast nearest point algorithms exist for these specific lattices.  Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.

@@

During the proof of strong consistency we made use of the fact that
\begin{equation}~\label{eq:SNVNunifmlln2}
\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
\end{equation}
almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here.  In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$. Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.  Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.

\begin{lemma} \label{lem:vn} $\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 2$.
 \end{lemma}
\begin{proof}
Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
 %\begin{align*}
\[ 
%\lambdabf[\rbf] = &\left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \right.\\ 
%&\dots, \; \frac{r_k}{N^{b+k}} - \frac{1}{2(k!)}, \; \dots \\
%&\left. \dots, \; \frac{r_m}{N^{b+m}} - \frac{1}{2(m!) }\right]
 \lambdabf[\rbf] = \left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \; \dots, \; \frac{r_m}{m!N^{b+m}} - \frac{1}{2(m!) }\right]
\]
 %\end{align*}
 for some constant $b>0$.  Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the zeroth coordinate, $\tfrac{1}{N^{b+1}}$ in the first coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate. Let
 \[
 B[\rbf]=\left\{  \xbf\in\reals^{m+1}  ; \frac{r_k}{N^{b+k}}\leq x_k + \frac{1}{2(k!)} < \frac{r_k + 1}{N^{b+k}} \right\}  .
 \]
and let $G$ be the finite set of grid points
\[
G = \left\{ \xbf \in \ints^{m+1} \mid x_k = 0,1,2\dots,N^{b+k}-1  \right\}.
\]
The total number of grid points is $|G| = N^{(m+1)(2b + m)/2}$, and the $B[\rbf]$ partition $B$, that is, $B = \cup_{\rbf \in G}B[\rbf]$.  Now 
\begin{align}
 \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
 &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
 \end{align}
From Lemma~\ref{lem:supVjk} it will follow that
\[
\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
\]
for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
\[
\sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\abs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])} < 2\frac{m+1}{N^b}.
\]
%that is, $D_N$ does not change much within any of the regions $B[\rbf]$ neighbouring the grid points.  
Combining these results with~\eqref{eq:gridanddensespace}, we obtain
\[
\prob\left(\sup_{\lambdabf \in B }\sabs{ D_N(\lambdabf)} > \frac{\epsilon}{2} +  \frac{2(m+1)}{N^b} \right) = O(e^{-c \epsilon^2 N}),
\]
and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
\end{proof}

@@

\begin{lemma}\label{lem:supVjk}
$\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 8$.
\end{lemma}
 \begin{proof}
Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
 \[
 Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - \expect \fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
 \]
 are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
 \[
  \prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
 \]
and so,
 \begin{align*}
 \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
&= 2 |G| e^{ - 8 \epsilon^2 N} = O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
\end{align*}
where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
\end{proof}

@@

\begin{lemma}\label{lem:boundedsquarefracparts}
$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all real numbers $x$ and $\delta$.
\end{lemma}
\begin{proof}
Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Also, for reasons of symmetry, we need only show that it holds when $\delta \geq 0$.  Now
\[
\fracpart{x + \delta}^2 - x^2 = \begin{cases}
2x\delta + \delta^2, & x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta) \\
2x(\delta-1) + (\delta-1)^2, & x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})
\end{cases}
\] 
But, when $x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta)$, 
\[
-1 \leq -1 + \delta \leq 2x + \delta < 1 - \delta \leq 1,
\]
and so
\[
-\delta \leq (-1 + \delta)\delta \leq (2x + \delta)\delta < (1-\delta)\delta \leq \delta.
\]
Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$ we  have $-\delta \leq 2x + \delta - 1 < \delta$, and consequently
\[
-\delta \leq -\delta(1-\delta) \leq (2x + \delta - 1)(1 - \delta) \leq \delta(1 - \delta) \leq \delta.
\]
%and consequenly
%\[
%-\delta \leq (2x + \delta - 1)(\delta - 1) \leq \delta.
%\]

@@

 \begin{lemma}\label{lem:supBVn} $\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}$ for all $N$.
 \end{lemma}
 \begin{proof}
%The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
For convenience put 
\[
b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \qquad \text{and} \qquad  a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
\] 
where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
where 
\[
\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
\] 
From Lemma~\ref{lem:boundedsquarefracparts} it follows that $-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|$,
and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all real numbers $x$. Now
\[
S_N(\lambdabf) - S_N(\lambdabf[\rbf]) =  \frac{1}{N}\sum_{n=1}^{N}\big( \fracpart{\Phi_n + b_n}^2 -  \fracpart{\Phi_n + a_n}^2 \big)
\]
and therefore $\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}$ for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
\[
\sabs{ V_N(\lambdabf) - V_N(\lambdabf[\rbf])} \leq \expect \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
\]
by Jensen's inequality.  Therefore, for all $\lambdabf \in B[\rbf]$,
 \begin{align*}
 \sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } &= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
&\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
&\leq 2\frac{m+1}{N^{b}},
\end{align*}
and the lemma follows because this bound is independent of $\rbf$.
%\[
%\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\sabs{ D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right) } \leq 2\frac{m+1}{N^{b}}.
%\]
 \end{proof}

@@

During the proof of asymptotic normality in Lemma~\ref{lem:Kconvfhalf} we made use of the following result regarding the function,
\[
G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
\]
where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.  To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.  The proof we will give holds for any nonegative integers $\ell$.

@@

\begin{align}
 \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \nonumber \\
